The \dealii Library, Version 9.0

### 2018-04-24 9.0/paper.tex
Write more about matrix-free, update references.

The major changes of this release are:
\begin{itemize}
\item Improved support for curved geometries;
\item Support for particle-in-cell methods;
\item Dedicated support for automatic and symbolic differentiation;
\item Interfaces to more external libraries and programs;
\item C++11 is now both required and used;
\item Support for GPU computations;
\item Support for face integrals and significant improvements of matrix-free framework;
\end{itemize}
These will all be discussed in more detail in the
following section. In addition, this release contains the following changes:
\begin{itemize}
\item \dealii has made extensive use of both the Clang-Tidy \cite{clang-tidy}
  and Coverity Scan \cite{coverity} static analysis tools for detecting bugs
  and other issues in the code. For example, around 260 issues were detected and
  fixed using the latter tool.

@@

\item New non-standard quadrature rules:
A number of non-standard, special-purpose quadrature rules have been
implemented. Among these are ones for
(i) truncating standard formulas to simplical domains (\texttt{QSimplex});
(ii) singular transformations of the unit cell to the unit simplex
    (\texttt{QDuffy});
(iii) composition of simplical quadrature rules to a combined rule on the
    unit cell (\texttt{QSplit});
and (iv) transformation of the unit square to polar coordinates
    (\texttt{QTrianglePolar}).
These quadrature rules greatly help when integrating singular
functions or on singular domains. They
are mainly used in Boundary Element Methods.

@@

\paragraph*{Assimp, the Open Asset Import Library.}
  Assimp \cite{assimp} can be used to read about 40 different 3D
  graphics formats. A subset of these formats can be now be read from
  within \dealii to generate two-dimensional meshes, possibly embedded in
  a three-dimensional space.

% \paragraph*{Gmsh.}
%   \marginpar{Do we actually provide any functionality with the GMSH
%   executable yet? If so, complete this part}
%   Gmsh \cite{geuzaine2009gmsh}
% The only thing we currently do is to define DEAL_II_GMSH_EXECUTABLE_PATH

\paragraph*{nanoflann, a library for building and querying
  $k$-d trees of datasets.} Operations such as finding the vertex or cell
  closest to a given evaluation point occur frequently in many applications
  that use unstructured meshes. While the naive algorithm is linear in the
  number of vertices or cells, many such operations can be made significantly
  faster by building a $k$-d tree data structure that recursively subdivides
  a $k$ dimensional space. The nanoflann library \cite{nanoflann} provides
  such a data structure and allows querying it, either for closest points
  (e.g., when finding the closest vertex) or for searching the points that
  fall within a radius of a target point. This functionality is now available
  via \dealii interfaces.

\paragraph*{ROL, a Rapid Optimization Library.}
  ROL \cite{ridzal2014rapid} is a package for large-scale optimization.
  \marginpar{Can we add a sentence or two about the ROL interface here?}

\paragraph*{ScaLAPACK, a parallel dense linear algebra library.}
  ScaLAPACK \cite{slug} provides block-cyclic matrix distribution over 2D
  process grid. The functionality and interface of our wrappers is similar
  to the LAPACK wrappers for serial dense linear algebra, namely
  matrix-matrix multiplication, Cholesky and LU factorizations,
  eigensolvers, SVD, pseudoinverses, and save/load operations using
  the HDF5 format. All of this functionality is available even in
  cases where the number of MPI processes does not exactly divide the
  number of tiles stored in the block-cyclic data distribution.

@@

\paragraph*{SUNDIALS, a SUite of Nonlinear and DIfferential/ALgebraic
  Equation Solvers.}
  Solving nonlinear algebraic and differential equations is both a
  common task and one that often requires sophisticated globalization
  algorithms for efficiency and reliability. SUNDIALS \cite{sundials}
  provides these in a widely used format, both sequentially and in parallel.

@@

@@

\marginpar{All: update gmsh reference}
\dealii can interface with many other libraries:
\begin{multicols}{3}
\begin{itemize}
\item ADOL-C \cite{adol-c}
\item ARPACK \cite{arpack}
\item Assimp \cite{assimp}
\item BLAS and LAPACK
\item Gmsh \cite{geuzaine2009gmsh}
\item GSL \cite{gsl2016}
\item HDF5 \cite{hdf5}
\item METIS \cite{karypis1998fast}
\item MUMPS \cite{ADE00,MUMPS:1,MUMPS:2,mumps-web-page}
\item muparser \cite{muparser-web-page}
\item nanoflann \cite{nanoflann}
\item NetCDF \cite{rew1990netcdf}
\item OpenCASCADE \cite{opencascade-web-page}
\item p4est \cite{p4est}
\item PETSc \cite{petsc-user-ref,petsc-web-page}
\item ROL \cite{ridzal2014rapid}
\item ScaLAPACK \cite{slug}
\item SLEPc \cite{Hernandez:2005:SSF}
\item SUNDIALS \cite{sundials}
\item TBB \cite{Rei07}
\item Trilinos \cite{trilinos,trilinos-web-page}
\item UMFPACK \cite{umfpack}
\end{itemize}
\end{multicols}
Please consider citing the appropriate references if you use interfaces to these
libraries.

>>>>

The major changes of this release are:
\begin{itemize}
\item Improved support for curved geometries;
\item Support for particle-in-cell methods;
\item Dedicated support for automatic and symbolic differentiation;
\item Interfaces to more external libraries and programs;
\item C++11 is now both required and used;
\item Support for GPU computations;
\item Support for face integrals and significant improvements of matrix-free framework;
\end{itemize}
These will all be discussed in more detail in the
following section. In addition, this release contains the following changes:
\begin{itemize}
\item \dealii has made extensive use of both the Clang-Tidy \cite{clang-tidy}
  and Coverity Scan \cite{coverity} static analysis tools for detecting bugs
  and other issues in the code. For example, around 260 issues were detected and
  fixed using the latter tool.
      %% > Is there somewhere we can cite for either of these tools? It would
      %% > be nice to recognize the department of homeland security for offering
      %% The contract with homeland security was completed in 2009.

@@

\item New non-standard quadrature rules:
A number of non-standard, special-purpose quadrature rules have been
implemented. Among these are ones for
(i) truncating standard formulas to simplical domains (\texttt{Qsimplex});
(ii) singular transformations of the unit cell to the unit simplex
    (\texttt{QDuffy});
(iii) composition of simplical quadrature rules to a combined rule on the
    unit cell (\texttt{QSplit});
and (iv) transformation of the unit square to polar coordinates
    (\texttt{QTrianglePolar}).
These quadrature rules greatly help when integrating singular
functions or on singular domains. They
are mainly used in Boundary Element Methods.

@@

\marginpar{Add references to all of these libraries.}

\paragraph*{ScaLAPACK, a parallel dense linear algebra library.}
  ScaLAPACK provides block-cyclic matrix distribution over 2D process
  grid. The functionality and interface of our wrappers is similar to
  the LAPACK wrappers for serial dense linear algebra, namely
  matrix-matrix multiplication, Cholesky and LU factorizations,
  eigensolvers, SVD, pseudoinverses, and save/load operations using
  the HDF5 format. All of this functionality is available even in
  cases where the number of MPI processes does not exactly divide the
  number of tiles stored in the block-cyclic data distribution.

@@

\paragraph*{SUNDIALS, a SUite of Nonlinear and DIfferential/ALgebraic
  Equation Solvers.}
  Solving nonlinear algebraic and differential equations is both a
  common task and one that often requires sophisticated globalization
  algorithms for efficiency and reliability. SUNDIALS provides these
  in a widely used format, both sequentially and in parallel.

@@

\paragraph*{ROL, a Rapid Optimization Library.}
ROL is a package for large-scale optimization.
\marginpar{Can we add a sentence or two about the ROL interface here?}

\paragraph*{nanoflann, a library for building and querying
  $k$-d trees of datasets.} Operations such as finding the vertex
or cell closest
to a given evaluation point occur frequently in many applications that
use unstructured meshes. While the naive algorithm is linear in the
number of vertices or cells, many such operations can be made
significantly faster by building a $k$-d tree data structure that
recursively subdivides a $k$ dimensional space. The nanoflann library
provides such a data structure and allows querying it, either for
closest points (e.g., when finding the closest vertex) or for
searching the points that fall within a radius of a target point. This
functionality is now available via \dealii interfaces.

\paragraph*{Assimp, the Open Asset Import Library.}
  Assimp can be used to read about 40 different 3D graphics formats. A
  subset of these formats can be now be read from within \dealii to
  generate two-dimensional meshes, possibly embedded in a
  three-dimensional space.

\paragraph*{Gmsh.}
  \marginpar{Do we actually provide any functionality with the GMSH
  executable yet? If so, complete this part}

@@

\marginpar{All: update gmsh reference}
\dealii can interface with many other libraries:
\begin{multicols}{3}
\begin{itemize}
\item ADOL-C \cite{adol-c}
\item ARPACK \cite{arpack}
\item Assimp \cite{assimp}
\item BLAS and LAPACK
\item Gmsh \cite{...}
\item GSL \cite{gsl2016}
\item HDF5 \cite{hdf5}
\item METIS \cite{karypis1998fast}
\item MUMPS \cite{ADE00,MUMPS:1,MUMPS:2,mumps-web-page}
\item muparser \cite{muparser-web-page}
\item nanoflann \cite{nanoflann}
\item NetCDF \cite{rew1990netcdf}
\item OpenCASCADE \cite{opencascade-web-page}
\item p4est \cite{p4est}
\item PETSc \cite{petsc-user-ref,petsc-web-page}
\item ROL \cite{ridzal2014rapid}
\item ScaLAPACK
\item SLEPc \cite{Hernandez:2005:SSF}
\item SUNDIALS \cite{sundials}
\item TBB \cite{Rei07}
\item Trilinos \cite{trilinos,trilinos-web-page}
\item UMFPACK \cite{umfpack}
\end{itemize}
\end{multicols}
Please consider citing the appropriate references if you use interfaces to these
libraries.
